maxent model avec xvalidation (year-based test training)

### 2019-04-24 src/maxent_alldata_avec_xvalidation.Rmd
update to all data

```{r}
train <- prback
trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))
```

@@

```{r train and test inc AIC for reg val}
aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
#betavallist <- cbind(aicetc$regularization)
betavallist <- 0.25
  timeslice <- "all_data"
  train <- prback
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  #for (h in 1:nrow(betavallist)){
   # betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist)))

@@

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/", timeslice, "/models/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
    bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/", timeslice, "/models/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/", timeslice, "/models/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

@@

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
      #}

write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", betaval, "_AICforMaxentReg.csv"), row.names = FALSE)

>>>>

```{r}
# train <- prback
# trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
# train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
# timeslice <- "all_data"
# betavallist <- 0.25
# mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))
```

@@

```{r train and test inc AIC for reg val}
aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betavallist <- cbind(aicetc$regularization)
timeslice <- "all_data"
train <- prback
trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
nfolds <- cbind(c(1:5))

  for (h in 1:nrow(betavallist)){
    betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/", timeslice, "/models"), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)))

@@

    #count number of non-zero coefficients, excluding the final four rows
    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)

for(f in 1:nrow(nfolds)){
  modn <- nfolds[f]-1
      lamf <- lambdas_list[f]
      sampf <- samppred_list[f]
      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
      coefile <- head(coefile, -4)
      coefile2 <- subset(coefile, V2 != 0)
      nocoeff <- nrow(coefile2)
      bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
      bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
      #natural log
      samplepred <- read.csv(paste0(sampf), header = TRUE)
      samplepred$nat_log <- log(samplepred$Raw.prediction)
      sumsamplepred <- sum(samplepred$nat_log)

      #no presences used in model (less those used for testing)

      notrainingpoints <- maxresults$X.Training.samples[f]
      testauc <- maxresults$Test.AUC[f]
      trainauc <- maxresults$Training.AUC[f]
      modno <- maxresults$Species[f]

      aicetc$train_AUC <- trainauc
      aicetc$test_AUC <- testauc
      aicetc$n <- notrainingpoints
      aicetc$coefficients <- nocoeff
      aicetc$sum.of.logs <- sumsamplepred

      aicetc$bic <- bic #not sure this is correct...
      aicetc$beta <- betamval
      aicetc$foldno <- modn

@@

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
    write.csv(aicetc, paste0("../output/maxent/", timeslice, "/models/", betamval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
      }

  # aa <- paste0("fold", modn, "_", betamval, "_", "AICforMaxentReg.csv")
    aicperbetalst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = paste0("^", betamval, ".*AICforMaxentReg.csv"), full.names = TRUE)
    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
    aicperbeta <- do.call(rbind, aicperbeta)
    aicbetamean <- colMeans(aicperbeta)
    aicperbeta <- rbind(aicperbeta, aicbetamean)
    aicperbeta$foldno[f+1] <- f+1
    write.csv(aicperbeta, paste0("../output/maxent/", timeslice, "/models/aicperbetasum_", betamval, ".csv"), row.names = FALSE)

    betaavglst <- list.files(paste0("../output/maxent/", timeslice, "/models/"), pattern = "aicperbetasum_.*csv", full.names = TRUE)
    bavg <- lapply(betaavglst, read.csv, header = TRUE)
    bavg <- do.call(rbind, bavg)
    bavg <- subset(bavg, foldno == f+1)
    write.csv(bavg, paste0("../output/maxent/", timeslice, "/models/avgbetalst.csv"), row.names = FALSE)
  }
